\maketitle

\section{Hybrid Merge-Insertion Sort Analysis}
\begin{enumerate}
\item Let $n$ be the input size / no. of elements to be sorted.
\item Divide $n$ elements into $\frac{n}{S}$ subarray of size $S$, the size of each subarray.
\item For each subarray of size $S$, perform insertion sort:
  \begin{itemize}
    \item Best case: $O(S)$
    \item Worst case: $O(S^2)$
    \item Average case: $O(S^2)$
  \end{itemize}
\item Since there are $\frac{n}{S}$ subarrays of size $S$, total cost of applying insertion sort as:
  \begin{itemize}
    \item Best case: $\frac{n}{S} \cdot S = O(n)$
    \item Worst case: $\frac{n}{S} \cdot S^2 = O(nS)$
    \item Average case: $\frac{n}{S} \cdot S^2 = O(nS)$
  \end{itemize}
   \pagebreak
\item Merge $\frac{n}{S}$ sorted subarray of size $S$ \\
  \begin{tikzpicture}[
      grow=up,
      level/.style={sibling distance=60mm/#1}]
    \node {$2^l S$} child foreach \n in {1,2} {
      node {$2^{l-1}S$} child foreach \m in {1,2} {
        node {$2^{l-2}S$} child  {
          node{$2S$} edge from parent[dashed] child foreach \i in {1,2} {
            node {$S$} edge from parent[solid]
          }
        }
      }
    };
  \end{tikzpicture}
  \begin{itemize}
    \item Each \texttt{merge()} doubles the subarray size. Suppose $l$ iterations of
      \texttt{merge()} has to be performed to recover input size $n$:
      \begin{equation}
        \begin{split}
          2^lS &= n \\
          2^l &= \frac{n}{s} \\
          l\lg{2} &= \lg{(\frac{n}{s})} \\
          l &= \log_2{(\frac{n}{s})}
        \end{split}
      \end{equation}
    \item Since the cost of \texttt{merge()} is $O(n)$, the worst case total cost of merging $M$ is:
      \begin{equation}
        \begin{split}
          M &= l \cdot n \\
            &= \log_2{(\frac{n}{s})} \cdot n \\
            &= O(\lg{(\frac{n}{s})} \cdot n)
        \end{split}
      \end{equation}
  \end{itemize}
  \item Combining insertion-sort \& merging in Hybrid Merge-Insertion sort, we have:
    \begin{itemize}
      \item Best Case: $O(n + \lg{(\frac{n}{s})} \cdot n)$
      \item Worst Case: $O(nS + \lg{(\frac{n}{s})} \cdot n)$
      \item Average Case: $O(nS + \lg{(\frac{n}{s})} \cdot n)$
    \end{itemize}
  \pagebreak
  \item To derive an exact equation for total no. of keys comparisons. 
    \begin{itemize}
      \item For Insertion Sort in the average case, $i$th iteration's no. of key comparisons could range $1 \to i$ with uniform probability $\frac{1}{i}$.
      \item Expected no. of key comparisons $= \frac{1}{i}\sum^i_{j=1}{j}$
      \item Since there are $S-1$ iterations, total no. of comparisons:
      \begin{equation}
        \sum^{S-1}_{i=1}\frac{1}{i}\sum^i_{j=1}{j} = \frac{1}{2}(\frac{(S-1)(S+2)}{2}) \approx \frac{1}{4}S^2
      \end{equation}
    \item  Total no. of comparisons for insertion sort = $\frac{1}{4}S^2 \cdot \frac{n}{S} = \frac{1}{4}nS$
      \item For Merge Sort, the no. of comparisons = $\lg{(\frac{n}{s})} \cdot n$
      \item Total no. of comparisons for hybrid sort = $\frac{1}{4}nS + \lg{(\frac{n}{s})} \cdot n$
    \end{itemize}
\end{enumerate}
